---
title: "hw-04-join-cheatsheet"
author: "Melanie Ang"
date: "October 5, 2017"
output: html_document
---

# Melanie's cheatsheet for dplyr::join featuring FISHES!

We are working with 2 data frames:  
* Canada_fishes: 7 common marine fishes found in Canadian waters
* Fish_habitat: Some fun facts about where you can find these fishes. Courtesy of [FishBase](www.fishbase.ca)

Let's begin by loading these dataframes. These csv files are found within this github folder.

```{r}
Canada_fishes <- "
Common name, Scientific name
Pink salmon, Oncorhynchus gorbuscha
Chum salmon, Oncorhynchus keta
Sockeye salmon, Oncorhynchus nerka
Chinook salmon, Oncorhynchus tshawytscha
Coho salmon, Oncorhynchus kisutch
Pacific halibut, Hippoglossus stenolepis
Pacific herring, Clupea pallasii pallasii
"
Canada_fishes <- read.csv("~/STAT545-hw-Ang-Melanie/hw-04/Canada_fishes.csv")

Fish_habitat <- "
Scientific name, Northern Latitude, Southern Latitude, Maximum depth
Oncorhynchus gorbuscha, 79N, 29N, 250m
Oncorhynchus keta, 67N, 24N, 250m
Oncorhynchus nerka, 72N, 42N, 250m
Oncorhynchus tshawytscha, 72N, 27N, 375m
Oncorhynchus kisutch, 72N, 22N, 250m
Hippoglossus stenolepis, 73N, 42N, 1200m
Gadus chalcogrammus, 68N, 34N, 400m
"

Fish_habitat <- read.csv("~/STAT545-hw-Ang-Melanie/hw-04/Fish_habitat.csv")

```

Lets take a look at these dataframes more closely:

```{r include = FALSE}
Canada_fishes_kable <- knitr::kable(Canada_fishes)
Fish_habitat_kable <- knitr::kable(Fish_habitat)
```

<table border = 1>
<tr>
<td valign="top">
  Canada fishes
  `r Canada_fishes_kable`
</td>
<td valign="top">
  publishers
  `r Fish_habitat_kable`
</td>
</tr>
</table>

As you can see they are both linked through the common column, Scientific name.


