---
title: "hw-03"
author: "Melanie Ang"
date: "September 27, 2017"
output: pdf_document
---

# Homework 3
In this homework, I experiment with dplyr (to manipulate data) and ggplot (to visualize data). Click [here](http://stat545.com/hw03_dplyr-and-more-ggplot2.html) to see the assignment on the STAT545 homework on web

### Step 1: Download packages
```{r packages, warning = FALSE, results='hide', message=FALSE}
library(gapminder) # our data set
library(tidyverse) # dplyr and ggplot2
```

### Task menu
I've picked four task from the selection on the [task section of the website](http://stat545.com/hw03_dplyr-and-more-ggplot2.html#task-menu) 

**1. Get the maximum and minimum of GDP per capita for all continents.**

```{r question 1 table}
gapminder %>%
  select(continent, gdpPercap) %>%  # subset only relevant columns so its neater
  group_by(continent) %>%           # group according to continents
  summarize(min_gdppercap=min(gdpPercap), 
            max_gdppercap=max(gdpPercap)) %>% 
  knitr::kable()                    # display in kable
```
```{r question 1 display plot}

# Taking the same function as above, I pipe it into a geom point plot
# geom_point() for continuous x and continuous y but I'm using this just to visualize the data
gapminder %>%
  select(continent, gdpPercap) %>%  # subset only relevant columns so its neater
  group_by(continent) %>%           # group according to continents
  summarize(min_gdppercap = min(gdpPercap),
            max_gdppercap = max(gdpPercap)) %>% # piping it into a ggplot
  ggplot(aes(x = continent)) +
    geom_point(aes(y=min_gdppercap), colour = "blue") +
    geom_point(aes(y=max_gdppercap), colour = "red") +
    labs(y = "GDP per Cap") +
    ggtitle("Minimum and maximum GDP per country")

# Now lets try a boxplot
# Boxplots are usually used for discrete x (like continent) and continuous y
gapminder %>%
  select(continent, gdpPercap) %>%  # subset only relevant columns so its neater
  ggplot(aes(x = continent, y = gdpPercap)) +
    geom_boxplot()

# Trying out a geom_violin() plot also for discrete x and continuous y
# And I add pretty colours aka customizations!
gapminder %>%
  select(continent, gdpPercap) %>%  # subset only relevant columns so its neater
  group_by(continent) %>%           # group according to continents
  summarize(min_gdppercap = min(gdpPercap),
            max_gdppercap = max(gdpPercap)) %>%
  ggplot() +
    geom_point(aes(x=min_gdppercap, y = max_gdppercap)) +
  facet_wrap(~continent)

    geom_point(aes(y=min_gdppercap), colour = "blue") +
    geom_point(aes(y=max_gdppercap), colour = "red") +
    labs(y = "GDP per Cap")
    
    
# display just minimum 
ggplot(p, aes(x=continent, y=min_gdppercap)) +
  geom_point()


ggplot(p, aes(x=continent)) +
  geom_point(aes(y=min_gdppercap), colour = "blue") +
  geom_point(aes(y=max_gdppercap), colour = "red") +
  labs(y = "GDP per Cap")

# boxplots? # try facetwrap
```

2. Look at the spread of GDP per capita within the continents  
Within spread, I chose to look at mean, standard deviation and median

```{r question 2 table}
gapminder %>%
  group_by(continent) %>% 
  summarize(mean_gdp = mean(gdpPercap), sd_gdp = sd(gdpPercap), median_gdp = median(gdpPercap))

```

```{r question 2 plot}
gapminder %>%
  group_by(continent) %>% 
  summarize(mean_gdp = mean(gdpPercap), sd_gdp = sd(gdpPercap), median_gdp = median(gdpPercap)) %>% 
  ggplot(aes(x = continent)) +                  # setting continent as base x axis
  geom_point(aes(y = mean_gdp), colour = "purple") + # change colour
  geom_point(aes(y = sd_gdp), colour = "turquoise", shape = 3) + # change shape
  geom_point(aes(y = median_gdp), colour = "orange", size = 5) + # change size
  labs(y = "GDP per cap")
```

3. How is life expectancy changing over time on different continents?

```{r question 3 table}
gapminder %>% 
  #arrange()
  group_by(continent) %>% 
  select(continent, lifeExp, year) %>% 
  knitr::kable()
# is there a better way to display this?
  
```

```{r question 3 plot}
gapminder %>% 
  group_by(continent) %>% 
  select(continent, lifeExp, year) %>%
  ggplot(aes(x=year, y=lifeExp)) +
  geom_line(aes(colour = continent)) 
```

4. Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.

```{r}

```

5. Make up your own! Between the dplyr coverage in class and the list above, I think you get the idea.



```{r}
gapminder %>% 
  filter(year %in% c("2000", "2002")) %>% 
  select(country, year, gdpPercap)
```

